\documentclass[12pt]{article}

\usepackage[margin=0.93in]{geometry}
\input{../../syllabi/preamble}

%%edit these!!
\newcommand{\coursedept}{Math}
\newcommand{\coursenumber}{340}
\newcommand{\professorname}{Adam Kapelner, PhD.}
\newcommand{\professorcontactinfo}{kapelner@qc.cuny.edu}
\newcommand{\professoroffice}{604 Kiely Hall}
\newcommand{\sixhundredsection}{640}
\newcommand{\coursenumbercrosslisted}{/ \sixhundredsection}
\newcommand{\semester}{Fall}
\newcommand{\numcredits}{4}
\newcommand{\lectimeandloc}{e.g. Mon and Wed 4:30 -- 6:20PM / KY258}
\newcommand{\requiredlabtimeandloc}{}
\newcommand{\tataofficehourtimeandloc}{} %TA / TA Office Hours / Loc & Kennly Weerasinghe / see \coursewebpagelink
\newcommand{\numtheoryhws}{6--9}
\newcommand{\lastdatetimetohandinhomeworks}{Dec 15 at noon}
\newcommand{\midtermonedateandlocation}{Thursday, October 7 on zoom during class time}
\newcommand{\midtermtwodateandlocation}{Thursday, November 11 on zoom during class time}
\newcommand{\finaldateandlocation}{TBD but on zoom}

%%%unchanged
\newcommand{\coursewebpageurl}{https://github.com/kapelner/QC_\coursedept_\coursenumber_\semester_\the\year}
\newcommand{\coursewebpagelink}{\href{\coursewebpageurl}{course homepage}}
\newcommand{\slackurl}{https://QC\coursedept\coursenumber\semester\the\year.slack.com/}
\newcommand{\slacklink}{\href{\slackurl}{slack}}

\input{../../syllabi/_header}



\section*{Course Overview}

MATH 340 / 640. Probability Theory for Data Science and Statistics. 4 hr.; 4 cr. Prereq.: 241. Coreq.: MATH 201 and 231. Convolutions, multivariate transformation of variables, the poisson process, beta, logistic, laplace, Weibull distributions, characteristic functions, central limit theorem, Cochran’s Theorem, the multivariate normal, chi-squared, T, F, Cauchy distributions, vector random variables, covariance matrices, multinomial, multivariate normal, order statistics, law of large numbers, famous inequalities, convergence in distribution and probability, laws of large numbers, Slutsky’s Theorem, optimization including stochastic gradient descent, Markov chains, Gibbs sampling. Probability computation using modern software. Special topics. Not open to students who are taking or who have received credit for MATH 340. Students cannot receive credit for more both: MATH 340, and 640. Fall, Spring. \pagebreak

\input{../../syllabi/_DSS_core}

Examining the above, we note that MATH 340 can be taken as a standalone course as a higher-level math elective. It can be thought of as advanced probability following MATH 241. This course is also recommended for students who are considering an actuarial career as this course provides great practice for Exam P. \pagebreak

\section*{Tentative Day-by-Day Schedule}

Lectures and their topics with rough time estimates per topic are below:

\input{_lecture_schedule}

\textbf{This is more of a typical mathematics theory course than the rest of the data science series.} But we will still attempt to keep our eye on developing ideas and concepts for helping to make decisions in the real world. Thus we may make limited use of computation using the \texttt{R} statistical language.

\subsection*{Prerequisites}

MATH 241 (basic probability), 201 (multivariable calculus) and 231 (linear algebra) or equivalents. I expect a 241 class that covers more or less what I cover in 241. See the course homepage for links under \qu{prerequisite review}. The multivariable calculus and linear algebra we will use will be light and I will try to review those concepts in class as we need them.

\section*{Course Materials}

\paragraph{Textbook:} Introduction to Probability Theory by Hoel, Port \& Stone. This book is out of print but you can buy it used on \href{https://www.amazon.com/Introduction-Probability-Theory-Paul-Hoel/dp/039504636X/ref=sr_1_1?ie=UTF8&qid=1503515517&sr=8-1&keywords=introduction+to+probability+theory+hoel}{Amazon} for $\approx$\$20, a reasonable price (as far as textbooks go). There is no excuse not to have this book. It is \textit{required}. However, I will not ususally be teaching \qu{from the book} --- most of the material in the class comes from the lecture notes. The textbook is a way to get ``another take'' on the material. The textbook covers about only half of the material done in class (yes, sometimes we will be following the textbook page by page). For the other half, you will have to make use of other resources.

\paragraph{Computer Software:} We will also be using \texttt{R} which is a free, open source statistical programming language and console. You can download it from: \url{http://cran.mirrors.hoobly.com/}. I do not expect you to do \textit{any} programming. I will be giving you \texttt{R} code to run and expect you to interpret the results based on concepts explained during the course.

\paragraph{Calculator:} You can use a TI-84, 85, 89 or any calculator which you wish. I strongly suggest you use \href{http://www.wolframalpha.com/}{Wolfram Alpha} and its smartphone app.

\input{../../syllabi/_the650section}

\input{../../syllabi/_use_of_slack}

\input{../../syllabi/_announcements_on_slack}

\input{../../syllabi/_standard_class_meetings}

%\input{../../syllabi/_jewish_holiday_reschedule}

%\input{../../syllabi/_zoom_policies}

%\input{../../syllabi/_lecture_upload}

\section*{Homework}

\input{../../syllabi/_theory_hws_text}

\input{../../syllabi/_theory_hws_submission_text}
\input{../../syllabi/_philosophy_hws}

\input{../../syllabi/_time_spent_hws}

\input{../../syllabi/_late_hw_policy}

\input{../../syllabi/_latex_hw_bonus_policy}

\input{../../syllabi/_hw_ec_policy}

\section*{Examinations}

\input{../../syllabi/_examination_text}

\input{../../syllabi/_standard_exam_schedule}

\subsection*{Exam Policies and Materials}

\input{../../syllabi/_examination_policies}

%\input{../../syllabi/_zoom_examination_policies}

\input{../../syllabi/_standard_cheat_sheet_policy}


\input{../../syllabi/_cheating_on_exams_and_missing_exams}
\input{../../syllabi/_special_services}

\input{../../syllabi/_class_participation}

%\input{../../syllabi/_zoom_attendance}

\input{../../syllabi/_standard_grading_and_grading_policy}

\input{../../syllabi/_advanced_course_grade_distribution}

\input{../../syllabi/_grade_checking_on_gradesly}

\input{../../syllabi/_auditing_policy}

\end{document}